---
title: "Time of Year Restrictions for Offshore Wind Development"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# load libraries
library(sf)
library(ggplot2)
library(cowplot)
library(boemWind)
source(here::here('R', 'utils.R'), echo = FALSE)

# Duke color scale
duke_colors = c('#7a0402', '#b01c02', '#d93b07', '#f26518', '#fc9c2d', '#f2c738', '#d0e835', '#a3fc3d', '#65fc6a', '#2deba2', '#24ced4', '#39a8f7', '#4678eb', '#4346ab', '#30123b')

# month order
month_order = c('October', 'November', 'December', 'January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September')

# month.abb
month.abb = substr(month_order, 1, 3)

# WEA order (roughly North to South)
wea_order = c('OCS-A 0486', 'OCS-A 0517', 'OCS-A 0501', 'OCS-A 0487', 'OCS-A 0500', 'OCS-A 0534', 'OCS-A 0520', 'OCS-A 0521', 'OCS-A 0522', 'OCS-A 0512', 'OCS-A 0544', 'OCS-A 0537', 'OCS-A 0538', 'OCS-A 0539', 'OCS-A 0549', 'OCS-A 0541', 'OCS-A 0542', 'OCS-A 0499', 'OCS-A 0498', 'OCS-A 0532', 'OCS-A 0482', 'OCS-A 0519', 'OCS-A 0490', 'OCS-A 0483', 'OCS-A 0508', 'OCS-A 0545', 'OCS-A 0546')
```

```{r data}
# see ./inst/scripts/prep_data.R

################
## NARW stuff ##
################

narw_weas= here::here('data', 'narw_dens_weas.rds') |>
  readRDS() |>
  dplyr::mutate(wea = factor(wea, levels = wea_order), month = factor(month, levels = month_order, labels = month.abb))
narw_10km = here::here('data', 'narw_dens_weas_10km.rds') |>
  readRDS() |>
  dplyr::mutate(wea_10km = factor(wea_10km , levels = wea_order), month = factor(month, levels = month_order, labels = month.abb))
narw_40km = here::here('data', 'narw_dens_weas_40km.rds') |>
  readRDS() |>
  dplyr::mutate(wea_40km  = factor(wea_40km , levels = wea_order), month = factor(month, levels = month_order, labels = month.abb))

####################
## NARW densities ##
####################

# coastwide
dens_coast = narw_weas |>
  sf::st_drop_geometry() |>
  dplyr::group_by(month) |> 
  dplyr::summarise(density = quantile(value, probs = 0.9, na.rm = TRUE)) |>
  dplyr::ungroup()

# region
dens_region = narw_weas |>
  dplyr::filter(!is.na(region)) |>
  sf::st_drop_geometry() |>
  dplyr::group_by(month, region) |> 
  dplyr::summarise(density = quantile(value, probs = 0.9, na.rm = TRUE)) |>
  dplyr::ungroup()
  
# wea
dens_wea = narw_weas |>
  dplyr::filter(!is.na(wea)) |>
  sf::st_drop_geometry() |>
  dplyr::group_by(month, wea) |> 
  dplyr::summarise(density = quantile(value, probs = 0.9, na.rm = TRUE)) |>
  dplyr::ungroup()
  
# 10km wea
dens_10km = narw_10km |>
  dplyr::filter(!is.na(wea_10km)) |>
  sf::st_drop_geometry() |>
  dplyr::group_by(month, wea_10km) |> 
  dplyr::summarise(density = quantile(value, probs = 0.9, na.rm = TRUE)) |>
  dplyr::ungroup()

############
## Shapes ##
############

# ctoy regions
ctoy_regions = here::here('data-raw', 'shapefiles', 'regions', 'CTOY_Investigation.shp') |>
  sf::st_read(quiet = TRUE) 

# BOEM active lease areas
active_weas = boem_wea_outlines |>
  dplyr::filter(LEASE_STAGE == 'Active' & !STATE %in% c('CA', ' Louisanna/Texas') & LEASE_TYPE != 'Easement' & LEASE_NUMBER_COMPANY != 'OCS-A 0506 - The Narragansett Electric Company') |>
  sf::st_transform(crs = sf::st_crs(ctoy_regions))
```

Quick look at the North Atlantic Right Whale (NARW) densities from MGEL, Duke University. Superimposed are the CTOY regions and active lease Wind Energy Areas (WEAs; although they are hard to see).

```{r narw-map, fig.width = 8, fig.height = 8, fig.align = 'left'}
# data for ggplot2
narw_gg = narw_weas |>
  dplyr::mutate(x = sf::st_coordinates(geometry)[, 1], y = sf::st_coordinates(geometry)[, 2]) |>
  sf::st_drop_geometry() |>
  dplyr::mutate(fill = cut(value * 100, breaks = c(6.3, 4, 2.5, 1.6, 1, 0.63, 0.40, 0.25, 0.16, 0.10, 0.063, 0.040, 0.025, 0.016, 0), include.lowest = TRUE))

# ggplot2
ggplot(data = narw_gg) + 
  geom_raster(aes(x = x, y = y, fill = fill)) + 
  geom_sf(data = ctoy_regions, fill = 'transparent', color = 'white') +
  geom_sf(data = active_weas, fill = 'transparent', color = 'white') +
  facet_wrap(~month) + 
  theme_cowplot() +
  scale_fill_manual(
    values = rev(duke_colors), 
    guide = guide_legend(reverse = TRUE),
    name = 'whales / 100km2') + 
  coord_sf(crs = sf::st_crs(narw_weas)) + 
  labs(x = NULL, y = NULL) + 
  theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold')) + 
  scale_x_continuous(breaks = c(-80, -70))
```

Now a quick look at the distribution of seasonal NARW densities by active lease WEAs (roughly ordered from north to south). The upper whisker of the boxplots denotes the 90th percentile.

```{r boxplot, fig.width = 8, fig.height = 50, fig.align = 'left'}
ggplot(data = narw_weas |> sf::st_drop_geometry() |> dplyr::filter(!is.na(wea)), aes(x = month, y = value * 100, fill = month)) +
  stat_summary(fun.data = boxplot_stat, geom = 'boxplot') + 
  facet_wrap(~wea, ncol = 1, scales = 'free_y', drop = TRUE) + 
  theme_cowplot() + 
  theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold')) + 
  labs(y = 'density (whales / 100km2)') + 
  guides(fill = 'none') 
  # + geom_hline(yintercept = 1, linetype = 'dashed') # floor((4 / 100) / 3.429904 * 100)
```

Now some line charts showing the 90th percentile of NARW densities extracted from various areas: 1) entire coast, 2) regions, 3) active lease WEAs, 4) active lease WEAs buffered by 10km, and 5) active lease WEAs buffered by 40 km (not yet implemented).

```{r trigger, fig.width = 8, fig.height = 50, fig.align = 'left'}
# sequence of trigger values
trigger = seq(0, 0.01, by = 0.001)[-1] # can't have the trigger be 0
j = trigger[1]

# loop through the weas and triggers
plt_list = vector('list', length(wea_order))
for (i in wea_order) {
    
    # filter data by wea
    wea_filt = dens_wea |>
      dplyr::filter(wea == i) |>
      dplyr::mutate(constraint = '0km')
    
    # filter data by wea bufferd by 10km
    wea_10km_filt = dens_10km |>
      dplyr::filter(wea_10km == i) |>
      dplyr::mutate(constraint = '10km') |>
      dplyr::rename(wea = wea_10km)
    
    # what region?
    wea_region = narw_weas |>
      dplyr::filter(wea == i) |>
      dplyr::distinct(region) |>
      dplyr::pull(region)
    
    # region
    region_filt = dens_region |>
      dplyr::filter(region == wea_region) |>
      dplyr::rename(constraint = region) |>
      dplyr::mutate(wea = i)
    
    # coastwide
    coast_filt = dens_coast |> 
      dplyr::mutate(constraint = 'Coastwide', wea = i)
    
    # combine
    data_gg = dplyr::bind_rows(wea_filt, wea_10km_filt, region_filt, coast_filt) |>
      dplyr::mutate(constraint = factor(constraint, levels = c('0km', '10km', wea_region, 'Coastwide'))) |>
      dplyr::mutate(ctoy = ifelse(density > j, TRUE, FALSE))

    # plot
    if (i == wea_order[length(wea_order)]) {
      plt = ggplot(data = data_gg, aes(x = month, y = density * 100, color = constraint, group = constraint)) + 
        geom_point(size = 3) +
        geom_line(linewidth = 1) + 
        theme_cowplot() + 
        facet_wrap(~wea) + 
        theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold'), legend.position = 'bottom', legend.title = element_blank()) + 
        labs(x = NULL, y = NULL) 
    } else {
      plt = ggplot(data = data_gg, aes(x = month, y = density * 100, color = constraint, group = constraint)) + 
        geom_point(size = 3) +
        geom_line(linewidth = 1) + 
        theme_cowplot() + 
        facet_wrap(~wea) + 
        theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold'), legend.position = 'bottom', legend.title = element_blank()) + 
        labs(x = NULL, y = NULL) 
    }

    # add to plot list
    plt_list[[which(wea_order %in% i)]] = plt
}

# plot
narw_plts = do.call(rbind, lapply(plt_list, ggplotGrob))
gridExtra::grid.arrange(narw_plts, left = grid::textGrob('density (whales / 100km2)', rot = 90, vjust = 0, gp = grid::gpar(fontsize = 14, fontface = 'bold')), bottom = grid::textGrob('month', vjust = 0, gp = grid::gpar(fontsize = 14, fontface = 'bold')))
```