---
title: "Time of Year Restrictions for Offshore Wind Development"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# load libraries
library(sf)
library(ggplot2)
library(cowplot)
library(boemWind)
source(here::here('R', 'utils.R'), echo = FALSE)
library(DT)

# Duke color scale
duke_colors = c('#7a0402', '#b01c02', '#d93b07', '#f26518', '#fc9c2d', '#f2c738', '#d0e835', '#a3fc3d', '#65fc6a', '#2deba2', '#24ced4', '#39a8f7', '#4678eb', '#4346ab', '#30123b')

# month order
month_order = c('October', 'November', 'December', 'January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September')

# month abb
month.abb = substr(month_order, 1, 3)

# WEA order (roughly North to South)
wea_order = c('OCS-A 0486', 'OCS-A 0517', 'OCS-A 0501', 'OCS-A 0487', 'OCS-A 0500', 'OCS-A 0534', 'OCS-A 0520', 'OCS-A 0521', 'OCS-A 0522', 'OCS-A 0512', 'OCS-A 0544', 'OCS-A 0537', 'OCS-A 0538', 'OCS-A 0539', 'OCS-A 0549', 'OCS-A 0541', 'OCS-A 0542', 'OCS-A 0499', 'OCS-A 0498', 'OCS-A 0532', 'OCS-A 0482', 'OCS-A 0519', 'OCS-A 0490', 'OCS-A 0483', 'OCS-A 0508', 'OCS-A 0545', 'OCS-A 0546')
```

```{r data}
# narw for ggplot2
narw_weas = here::here('data', 'narw_dens_weas.rds') |>
  readr::read_rds() |>
  dplyr::mutate(month = factor(month, levels = month_order, labels = month.abb),
                wea = factor(wea, levels = wea_order))
  
# narw for comparison
narw_dens = here::here('data', 'ctoy_density_final.csv') |>
  readr::read_csv() |>
  dplyr::mutate(month = factor(month, levels = month_order, labels = month.abb),
                wea = factor(wea, levels = wea_order))

# narw for trigger
narw_trigger = here::here('data', 'ctoy_density_trigger.csv') |>
  readr::read_csv() |>
  dplyr::mutate(month = factor(month, levels = month_order, labels = month.abb),
                wea = factor(wea, levels = wea_order))

# ctoy regions
ctoy_regions = here::here('data-raw', 'shapefiles', 'regions', 'CTOY_Investigation.shp') |>
  sf::st_read(quiet = TRUE) 

# BOEM active lease areas
active_weas = boem_wea_outlines |>
  dplyr::filter(LEASE_STAGE == 'Active' & !STATE %in% c('CA', ' Louisanna/Texas') & LEASE_TYPE != 'Easement' & LEASE_NUMBER_COMPANY != 'OCS-A 0506 - The Narragansett Electric Company') |>
  sf::st_transform(crs = sf::st_crs(ctoy_regions))
```

# Approach

Habitat-based North Atlantic right whale (NARW) density models for the U.S. Atlantic were downloaded from [OBIS-SEAMAP](https://seamap.env.duke.edu/models/Duke/EC/). The seasonal raster images were then deconstructed into data tables where each row represents a single grid cell, with columns representing attributes of that grid cell (e.g., cell id, cell centroid, cell area, month, density value). We then assigned each grid cell to several, nested areas by spatial intersection between the grid cell centroid and the considered areas polygons. All spatial layers used a bespoke Albers Equal Area coordinate reference system developed by the Marine Geospatial Ecology Lab, Duke University (CetMap_NA). Below are maps of the seasonal NARW density data, with superimposed white outlines that denote some of the considered areas.

```{r narw-map, fig.width = 8, fig.height = 8, fig.align = 'left'}
# data for ggplot2
narw_gg = narw_weas |>
  dplyr::mutate(x = sf::st_coordinates(geometry)[, 1], y = sf::st_coordinates(geometry)[, 2]) |>
  sf::st_drop_geometry() |>
  dplyr::mutate(fill = cut(value * 100, breaks = c(6.3, 4, 2.5, 1.6, 1, 0.63, 0.40, 0.25, 0.16, 0.10, 0.063, 0.040, 0.025, 0.016, 0), include.lowest = TRUE))

# ggplot2
ggplot(data = narw_gg) + 
  geom_raster(aes(x = x, y = y, fill = fill)) + 
  geom_sf(data = ctoy_regions, fill = 'transparent', color = 'white') +
  geom_sf(data = active_weas, fill = 'transparent', color = 'white') +
  facet_wrap(~month) + 
  theme_cowplot() +
  scale_fill_manual(
    values = rev(duke_colors), 
    guide = guide_legend(reverse = TRUE),
    name = 'whales / 100km2') + 
  coord_sf(crs = sf::st_crs(narw_weas)) + 
  labs(x = NULL, y = NULL) + 
  theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold')) + 
  scale_x_continuous(breaks = c(-80, -70))
```

In the following plots, we'll be summarizing the seasonal NARW density data by active lease Wind Energy Areas (WEAs) that are roughly ordered from north to south. A look up table that relates the lease number for active lease WEAs to their corresponding company and state is provided below.

```{r lookup}
# data for DT
data_dt = active_weas |> 
  dplyr::mutate(LEASE_NUMBER = factor(LEASE_NUMBER, levels = wea_order)) |>
  dplyr::arrange(LEASE_NUMBER) |>
  dplyr::filter(!is.na(LEASE_NUMBER)) |>
  sf::st_drop_geometry() |>
  dplyr::rename(Lease = LEASE_NUMBER, Company = COMPANY, State = STATE) |>
  dplyr::select(Lease, Company, State)
DT::datatable(data_dt, options = list(iDisplayLength = 50))
```

Below are boxplots of the seasonal NARW densities by active lease WEAs (again, roughly ordered from north to south). The boxplots were drawn such that the lower whisker denotes the 10th percentile, the lower hinge represents the 25th percentile, the median, the upper hinge represents the 75th percentile, and the upper whisker represents the 90th percentile.

```{r boxplot, fig.width = 8, fig.height = 50, fig.align = 'left'}
ggplot(data = narw_weas |> sf::st_drop_geometry() |> dplyr::filter(!is.na(wea)), aes(x = month, y = value * 100, fill = month)) +
  stat_summary(fun.data = boxplot_stat, geom = 'boxplot') + 
  facet_wrap(~wea, ncol = 1, scales = 'free_y') + 
  theme_cowplot() + 
  theme(strip.background = element_blank(), strip.text = element_text(hjust = 0, face = 'bold')) + 
  labs(y = 'density (whales / 100km2)') + 
  guides(fill = 'none') 
  # + geom_hline(yintercept = 1, linetype = 'dashed') # floor((4 / 100) / 3.429904 * 100)
```

Another useful summary are line charts showing the 90th percentile of the seasonal NARW densities extracted from various areas: 1) entire coast, 2) regions, 3) active lease WEAs, 4) active lease WEAs buffered by 10km, and 5) active lease WEAs buffered by 40 km.

```{r comparison, fig.width = 8, fig.height = 50, fig.align = 'left'}
# ggplot2
ggplot(data = narw_dens |> dplyr::mutate(constraint = factor(constraint, levels = c('0km', '10km', '40km', 'Southern New England', 'Mid-Atlantic', 'Southeast', 'Coastwide'))), aes(x = month, y = density * 100, color = constraint, group = constraint)) + 
    geom_point(size = 3) + 
    geom_line(linewidth = 1) +
    theme_cowplot() + 
    facet_wrap(~wea, ncol = 1, scales = 'free_y') + 
    theme(strip.background = element_blank(), 
          strip.text = element_text(hjust = 0, face = 'bold'), 
          legend.title = element_blank(),
          legend.position = 'bottom') + 
    labs(x = 'month', y = 'density (whales / 100km2)') 
```

## Case Study of Potential Thresholds for OCS-A 0501 - Vineyard Wind 1 LLC

Choosing a threshold for when to temporally restrict offshore wind farm (OWF) construction is complicated. Potential thresholds could include both relative and absolute measures. Relative measures can assess 'habitat preference', comparing a value extracted from a WEA to a value at the regional or coastwide scale. Relative thresholds over 1 would indicate that NARW density is higher inside the WEA relative to the region or the entire coast (or some other reference), potentially suggesting preferential use of the WEA by NARWs.

```{r triggerm, fig.width = 8.13, fig.height = 5.48, fig.align = 'left'}
# data for relative trigger
case_study = narw_trigger |>
  dplyr::filter(wea == 'OCS-A 0501') |>
  dplyr::arrange(month) |>
  dplyr::mutate(trigger_v1 = wea_0km / region,
                trigger_v2 = wea_0km / coastwide,
                trigger_v3 = wea_10km / region,
                trigger_v4 = wea_10km / coastwide,
                trigger_v5 = wea_40km / region,
                trigger_v6 = wea_40km / coastwide)

# wide to long
case_long = case_study |> 
  dplyr::select(-c(wea_0km:coastwide)) |>
  tidyr::pivot_longer(cols = trigger_v1:trigger_v6, names_to = 'trigger', values_to = 'value') |>
  dplyr::mutate(reference = dplyr::case_when(trigger %in% c('trigger_v1', 'trigger_v3', 'trigger_v5') ~ 'region', 
                                             trigger %in% c('trigger_v2', 'trigger_v4', 'trigger_v6') ~ 'coastwide'),
                wea = dplyr::case_when(trigger %in% c('trigger_v1', 'trigger_v2') ~ '0km', 
                                       trigger %in% c('trigger_v3', 'trigger_v4') ~ '10km', 
                                       trigger %in% c('trigger_v5', 'trigger_v6') ~ '40km'))

# plot it out
plt_rel = ggplot(data = case_long, aes(x = month, y = value, color = wea, group = trigger)) + 
  geom_point(size = 3) + 
  geom_line(linewidth = 1, aes(linetype = reference)) +
  theme_cowplot() + 
  geom_hline(yintercept = 1, linetype = 'dashed') + 
  labs(title = 'relative', y = 'wea / reference')

# data for absolute trigger
abs_trigger = narw_dens |> 
  dplyr::mutate(constraint = factor(constraint, levels = c('0km', '10km', '40km', 'Southern New England', 'Mid-Atlantic', 'Southeast', 'Coastwide'))) |>
  dplyr::filter(wea == 'OCS-A 0501')

# plot it out
plt_abs = ggplot(data = abs_trigger, aes(x = month, y = density * 100, color = constraint, group = constraint)) + 
  geom_point(size = 3) + 
  geom_line(linewidth = 1) + 
  theme_cowplot() + 
  theme(strip.background = element_blank(), 
        strip.text = element_text(hjust = 0, face = 'bold'), 
        legend.title = element_blank()) + 
  labs(title = 'absolute', y = 'density (whales / 100km2)') + 
  geom_hline(yintercept = 1, linetype = 'dashed')

# combine plots
plt = do.call(rbind, lapply(list(plt_rel, plt_abs), ggplotGrob))

# plot
gridExtra::grid.arrange(plt)
```